---
title: "Final Project"
author: "Shanie Talor, Arif Abd Aziz, Jon Girolamo"
date: '2023-06-02'
output: pdf_document
---

(We don't have to polish this RMD in terms of labels or descriptions. Final presentation will be in powerpoint. We'll still submit this as the source code though so we should still be commenting what we're doing with code) Jon

Load Data

```{r}
# Read csv
gold <- read.csv("gold.csv", sep = ",", header = TRUE)

# Omit all rows that have NA vals
gold <- na.omit(gold)

# Add Binary Classifier Variable
gold$Binary.PercChange <- ifelse(gold$PercChangeForc > 0.1, 1, 0)

# Hurdle rate for Binary CLassifier is set at 0.1%, so that edge cases that are
# very close to 0% (0%:0.0999%) are classified as non-investment opportunities

```

Variable Selection

```{r}
# Run Boruta  algorithm on clean data
library(Boruta)
library(randomForest)
boruta.gold <- Boruta(PercChangeForc~.-Date -Average.Price -Oil -PPIJewelry -Binary.PercChange, data = gold, doTrace = 2, 
                       randomForest = TRUE)

plot(boruta.gold)

```

Create Cross Validation Folds (Regression)

```{r}
library(caret)
set.seed(123)
folds <- createFolds(gold$PercChangeForc, k = 5, returnTrain = TRUE)
```

```{r}

```

# Regression

### OLS
```{r}

```

### Random Forest + Bagging
```{r}

```

### Boosting
```{r}

```


### Neural Network
```{r}

```


--------------------------------------------------------------------------------

# Classification

### Support Vector Machine
```{r}

```

### Neural Network
```{r}

```









